\textbf{NBA Party}

### 2017-06-01 init_model.rnw
Added table and plot showing effect of points admitted as a function of party.

\subsection{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
<<echo=F>>=
lines[, list(date = as.Date(date), team, opponent, line, score)]
@

@@

<<echo=F>>=
### Merge lines data with BLS data.
lines <- merge(lines, musicians,
               by.x = c('season', 'last.game.loc'),
               by.y = c('season', 'team'), all.x = T)

setkey(lines, season, team, date)

## lines[, list(date, team, opponent, location, 
##              line, score, outcome, nhours.lgame, 
##              travel.dist, avg.age, nmusicians)]

@@

HOURS <- 24
lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians), 0)]

@@

lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
data <- merge(lines, espn,
              by.x = c('team', 'season', 'simple.date'),
              by.y = c('team', 'season', 'date'), all.x = T)

@@

# Since we sum by location-date we double count.
changes <- data[, list(ttl.chg.pos = sum(chg.pos) / 2), 
            keyby = list(season, location, date)]
lines <- merge(lines, changes, by = c('season', 'location', 'date'))

setkey(lines, season, team, date)
lines[, lag.chg.pos := c(NA, lag(ttl.chg.pos)[1:.N-1]), by = list(season, team)]

@@

<<echo=F>>=
lines[, list(date = simple.date, location = substr(location, 1, 7), 
             team = substr(team, 1, 7), opponent = substr(opponent, 1, 7), 
             line, score, outcome, 
             avg.age, last.game.loc, nhours.lgame, travel.dist, party, lag.chg.pos)]
@

@@

lines[, ew := sin(travel.dir %% 360 / 360 * 2 * pi)]
#plot(lines$travel.dir, lines$ew)

m <- glm(outcome == 'W' ~ party + lag.chg.pos + travel.dist*ew + 
             nhours.lgame + I(hour(date)),
         data = lines[season < 2017], family = 'binomial',
         na.action = 'na.exclude')

summary(m)

# stargazer(m, style = 'qje', single.row = T)
@

@@

HOURS <- 36
m <- glm(outcome == 'W' ~ party, 
         data = lines[season < 2016 & last.game.loc != team])

### We're only interested in making predictions when teams party.
p <- predict(m, newdata = lines[season == 2016 & nhours.lgame <= HOURS & 
                                last.game.loc != team], type = 'response')

idx <- which(p < 0.476)
out <- table(lines[season == 2016 & nhours.lgame <= HOURS & last.game.loc != team][idx, outcome])

@@

table(round(p), lines[season == 2016 & nhours.lgame <= HOURS & 
                      last.game.loc != team, outcome])

@@

<<echo=1>>=
source('functions.r')
m <- glm(outcome == 'W' ~ party,
         data = lines[season < 2016 & last.game.loc != team])

p <- predict(m, newdata = lines[season == 2016 & nhours.lgame <= HOURS & 
                                last.game.loc != team], type = 'response')

table(round(p), lines[season == 2016 & nhours.lgame <= HOURS & 
                      last.game.loc != team, outcome])

@@

<<echo=F>>=
lines[, game.time := hour(date)]

m <- lm(demeaned.reb.per.min ~ party + lag.chg.pos + travel.dist + nhours.lgame,
   data = data[last.game.loc != team] %>% na.omit)

#stargazer(m, style = 'qje', single.row = T)
@

@@

<<echo=-1>>=
load(file = 'tmp_data/covers_lines_mlb.RData')
lines[, party := ifelse(team != location & weekend == 1, log(nmusicians), 0)]

m <- glm(I(team.score > opponent.score) ~ party + odds + I(team == location) + 
             ndays.lgame + travel.dist + weekend,
         data = lines, family = 'binomial')
summary(m)
# stargazer(m, style = 'qje', single.row = T)
@

@@

<<echo=2:4>>= 
m <- glm(I(team.score > opponent.score) ~ party + odds,
         data = lines[year(date) < 2016], family = 'binomial')
p <- predict(m, newdata = lines[year(date) == 2016], type = 'response')

idx <- which(p < .45)
p1 <- mapply(Predict, prediction = p[idx], 
             vegas = lines[year(date) == 2016, odds][idx],
             MoreArgs=list(threshold = 0.01))
b <- mapply(Bet, our.prediction = p1,
            actual.outcome = lines[year(date) == 2016, team.score > opponent.score][idx],
            odds = lines[year(date) == 2016, odds][idx], SIMPLIFY = T)

sum(b, na.rm = T)

### 
betting.days <- lines[year(date) == 2016][idx, date]
running.prof <- cumsum(ifelse(is.na(b), 0, b))
plot(x = 1:length(b), y = running.prof, 
     main = 'Profits over the 2016 season', 
     xlab = 'Game Number', ylab = 'Profit', type = 'l')
abline(a = 0, b = 0, lty = 3)
points(x = which.min(running.prof), y = min(running.prof), type = 'p')
text(x = which.min(running.prof), y = min(running.prof), labels = min(running.prof), pos = 4)

### the following plot shows sufficient deviation from house odds to motivate
### an optimized betting strategy.
hist(p- lines[year(date) == 2016, odds])

tholds <- seq(.01, .1, by = .005)
tmp <- sapply(tholds, function(t) {
    p1 <- mapply(Predict, prediction = p[idx], 
                 vegas = lines[year(date) == 2016, odds][idx],
                 MoreArgs=list(threshold = t))
    b <- mapply(Bet, our.prediction = p1,
                actual.outcome = lines[year(date) == 2016, team.score > opponent.score][idx],
                odds = lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
    return(c(sum(b, na.rm = T) / sum(!is.na(b)), sum(!is.na(b)))) # Avg profit per game.
})

@@

If we were to bet one hundred dollars on each game in 2016,
we would earn a profit of about \Sexpr{format(round(sum(b, na.rm=T), -1), big.mark = ',')} dollars
over the course of \Sexpr{sum(!is.na(b))} games.

>>>>

\subsection{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
<<echo=F>>=
nba.lines <- lines
nba.lines[, list(date = as.Date(date), team, opponent, line, score)]
@

@@

<<echo=F>>=
### Merge lines data with BLS data.
nba.lines <- merge(nba.lines, musicians,
               by.x = c('season', 'last.game.loc'),
               by.y = c('season', 'team'), all.x = T)

setkey(nba.lines, season, team, date)

### Exclude playoff games.
nba.lines <- nba.lines[, head(.SD, n=82), by=list(season,team)]

## nba.lines[, list(date, team, opponent, location, 
##              line, score, outcome, nhours.lgame, 
##              travel.dist, avg.age, nmusicians)]

@@

HOURS <- 24
nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians), 0)]

@@

nba.lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
data <- merge(nba.lines, espn,
              by.x = c('team', 'season', 'simple.date'),
              by.y = c('team', 'season', 'date'), all.x = T)

@@

# Since we sum by location-date we double count.
changes <- data[, list(ttl.chg.pos = sum(chg.pos) / 2), 
            keyby = list(season, location, date)]
nba.lines <- merge(nba.lines, changes, by = c('season', 'location', 'date'))

setkey(nba.lines, season, team, date)
nba.lines[, lag.chg.pos := c(NA, lag(ttl.chg.pos)[1:.N-1]), by = list(season, team)]

@@

<<echo=F>>=
nba.lines[, list(date = simple.date, location = substr(location, 1, 7), 
             team = substr(team, 1, 7), opponent = substr(opponent, 1, 7), 
             line, score, outcome, 
             avg.age, last.game.loc, nhours.lgame, travel.dist, party, lag.chg.pos)]
@

@@

nba.lines[, ew := sin(travel.dir %% 360 / 360 * 2 * pi)]
#plot(nba.lines$travel.dir, nba.lines$ew)

nba.lines[, party.discrete := grepl("(los angeles)|(new york)|(brooklyn)", last.game.loc) & nhours.lgame <= 24]

m0 <- glm(outcome == 'W' ~ party.discrete + lag.chg.pos + log(travel.dist+1)*ew + 
             nhours.lgame + I(hour(date)) + I(team==location),
         data = nba.lines[season < 2017], family = 'binomial',
         na.action = 'na.exclude')

m1 <- glm(outcome == 'W' ~ party + lag.chg.pos + log(travel.dist+1)*ew + 
             nhours.lgame + I(hour(date)) + I(team==location),
         data = nba.lines[season < 2017], family = 'binomial',
         na.action = 'na.exclude')

summary(m1)

## nba.lines[,.N/2,by=list(season)]
## stargazer(m)

@@

HOURS <- 36
m <- glm(outcome == 'W' ~ party, 
         data = nba.lines[season < 2016 & last.game.loc != team])

### We're only interested in making predictions when teams party.
p <- predict(m, newdata = nba.lines[season == 2016 & nhours.lgame <= HOURS & 
                                last.game.loc != team], type = 'response')

idx <- which(p < 0.476)
out <- table(nba.lines[season == 2016 & nhours.lgame <= HOURS & last.game.loc != team][idx, outcome])

@@

table(round(p), nba.lines[season == 2016 & nhours.lgame <= HOURS & 
                      last.game.loc != team, outcome])

@@

<<echo=1>>=
source('functions.r')
m <- glm(outcome == 'W' ~ party,
         data = nba.lines[season < 2016 & last.game.loc != team])

p <- predict(m, newdata = nba.lines[season == 2016 & nhours.lgame <= HOURS & 
                                last.game.loc != team], type = 'response')

table(round(p), nba.lines[season == 2016 & nhours.lgame <= HOURS & 
                      last.game.loc != team, outcome])

@@

<<echo=F>>=
nba.lines[, game.time := hour(date)]

m2 <- lm(demeaned.reb.per.min ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame,
   data = data[last.game.loc != team] %>% na.omit)
summary(m2)

lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame, data=data[last.game.loc!=team] %>% na.omit) %>% summary

### Points  allowed as afunction of last game location (by party)
nba.lines[, demeaned.pts.admitted := team.pts.admitted - mean(team.pts.admitted), by = list(season, team)]
nba.lines[, mean(demeaned.pts.admitted), by = list(last.game.loc, party.discrete)][order(V1)]
tmp <- nba.lines[, list(demeaned.pts.admitted = mean(demeaned.pts.admitted)), 
                 by = list(last.game.loc, party)][order(demeaned.pts.admitted)]
plot(x=jitter(tmp$party),y=tmp$demeaned.pts.admitted)
abline(lm(demeaned.pts.admitted ~ party, data=tmp))
#stargazer(m, style = 'qje', single.row = T)
## stargazer(m)

@@

<<echo=-1>>=
load(file = 'tmp_data/covers_lines_mlb.RData')
mlb.lines <- lines
mlb.lines[, party := ifelse(team != location & weekend == 1, log(nmusicians), 0)]

setkey(mlb.lines, season, team, date)
mlb.lines <- mlb.lines[, head(.SD, n=162),by=list(season,team,date)] #Exclude playoff games

m3 <- glm(I(team.score > opponent.score) ~ party + odds + I(team == location) + 
             ndays.lgame + log(travel.dist+1) + weekend,
         data = mlb.lines, family = 'binomial')
summary(m3)
#stargazer(m0, m1, m2, m3, single.row = T)
@

@@

<<echo=2:4>>=

pdf('mlb_bets_by_season.pdf')
par(mfrow = c(3,2))
b <- lapply(2012:2016, function(yr) {
    m <- glm(I(team.score > opponent.score) ~ party + odds,
             data = mlb.lines[year(date) < yr], family = 'binomial')
    p <- predict(m, newdata = mlb.lines[year(date) == yr], type = 'response')
    idx <- which(p < .45)
    p1 <- mapply(Predict, prediction = p[idx], 
                 vegas = mlb.lines[year(date) == yr, odds][idx],
                 MoreArgs=list(threshold = 0.01))
    b <- mapply(Bet, our.prediction = p1,
                actual.outcome = mlb.lines[year(date) == yr, team.score > opponent.score][idx],
                odds = mlb.lines[year(date) == yr, odds][idx], SIMPLIFY = T)
    cat(paste0(sum(b, na.rm = T), '\n'))
    ### 
#    pdf(paste0('mlb_bets_', yr, '.pdf'))
    betting.days <- mlb.lines[year(date) == yr][idx, date]
    running.prof <- cumsum(ifelse(is.na(b), 0, b))
    plot(x = 1:length(b), y = running.prof, 
         main = paste0('Profits over the ', yr, ' season\n',
                       'Training on 2011-', yr-1, ' data'),
         xlab = 'Game Number', ylab = 'Profit', type = 'l')
    abline(a = 0, b = 0, lty = 3)
    points(x = which.min(running.prof), y = min(running.prof), type = 'p')
    text(x = which.min(running.prof), y = min(running.prof), 
         labels = round(min(running.prof)), pos = 4)
    return(b)
#    dev.off()
})
dev.off()

\includegraphics{mlb_bets_by_season}

If we were to bet one hundred dollars on each game in 2016,
we would earn a profit of about \Sexpr{format(round(sum(b[[length(b)]], na.rm=T), -1), big.mark = ',')} dollars
over the course of \Sexpr{sum(!is.na(b))} games.

m <- glm(I(team.score > opponent.score) ~ party + odds,
         data = mlb.lines[year(date) < 2016], family = 'binomial')
p <- predict(m, newdata = mlb.lines[year(date) == 2016], type = 'response')
idx <- which(p < .45)
p1 <- mapply(Predict, prediction = p[idx], 
             vegas = mlb.lines[year(date) == 2016, odds][idx],
             MoreArgs=list(threshold = 0.01))
b <- mapply(Bet, our.prediction = p1,
                actual.outcome = mlb.lines[year(date) == 2016, team.score > opponent.score][idx],
            odds = mlb.lines[year(date) == 2016, odds][idx], SIMPLIFY = T)

### the following plot shows sufficient deviation from house odds to motivate
### an optimized betting strategy.
hist(p- mlb.lines[year(date) == 2016, odds])

tholds <- seq(.01, .1, by = .005)
tmp <- sapply(tholds, function(t) {
    p1 <- mapply(Predict, prediction = p[idx], 
                 vegas = mlb.lines[year(date) == 2016, odds][idx],
                 MoreArgs=list(threshold = t))
    b <- mapply(Bet, our.prediction = p1,
                actual.outcome = mlb.lines[year(date) == 2016, team.score > opponent.score][idx],
                odds = mlb.lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
    return(c(sum(b, na.rm = T) / sum(!is.na(b)), sum(!is.na(b)))) # Avg profit per game.
})

@@
